Out-of-Sample Comparisons of Overfit Models

### 2014-02-25 paper.tex
Change font to bitstream charter
Remove small caps

\documentclass[11pt]{article}
\usepackage{tikz,setspace}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage[round]{natbib}
\usepackage[margin=1in]{geometry}
\onehalfspacing
\input{macros}

@@

\title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Helle Bunzel, Marjorie Flavin,
    Nir Jaimovich, Lutz Kilian, Ivana Komunjer, Michael McCracken,
    Seth Pruitt, Ross Starr, Jim Stock, Yixiao Sun, Allan Timmermann,
    Hal White, several anonymous referees, participants at the 2010
    Midwest Economics Association Annual Meetings, the 2010
    International Symposium on Forecasting, the 2010 Joint Statistical
    Meetings, the 2011 \textsc{nber}/\textsc{nsf} Time-Series
    Conference, and in many seminars at \textsc{ucsd}, and espeically
    Graham Elliott for their valuable suggestions, feedback and advice
    in writing this paper.  I would also like to thank Amit Goyal for
    providing computer code and data for his 2008 \textsc{rfs} paper
    with Ivo Welch \citep{GoW:08}.} \\ Iowa State University}

@@

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Our analysis gives a theoretical motivation for using
  \oos\ comparisons: the \dmw\ \oos\ test allows a forecaster to
  conduct inference about the expected future accuracy of his or her
  models when one or both is overfit.  We show analytically and
  through Monte Carlo that standard full-sample test statistics can
  not test hypotheses about this performance.  Our paper also shows
  that popular test and training sample sizes may give misleading
  results if researchers are concerned about overfit.  We show that
  $P^2/T$ must converge to zero for the \dmw\ test to give valid
  inference about the expected forecast accuracy, otherwise the test
  measures the accuracy of the estimates constructed using only the
  training sample.  In empirical research, $P$ is typically much
  larger than this.  Our simulations indicate that using large values
  of $P$ with the \dmw\ test gives undersized tests with low power, so
  this practice may favor simple benchmark models too much.

\noindent \textsc{jel} Classification: C12, C22, C52, C53

@@

\section{Introduction}\label{sec:introduction}
Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use \oos\ instead of in-sample
comparisons: the \dmw\ \oos\ test\footnote{In this paper, we will
  refer to the basic \oos\ $t$-test studied by \citet{DiM:95} and
  \citet{Wes:96} as the \dmw\ test.} allows a forecaster to conduct
inference about the expected future accuracy of his or her models when
one or both is overfit.  We show analytically and through Monte Carlo
that standard full-sample test statistics can not test hypotheses
about this performance.

Our paper also shows that popular test and training sample sizes may
give misleading results if researchers are concerned about overfit.
We show that $P^2/T$ must converge to zero for the \dmw\ test to give
valid inference about the expected forecast accuracy, otherwise the
test measures the accuracy of the estimates constructed using only the
training sample.  In empirical research, $P$ is typically much larger
than this.  Our simulations indicate that using large values of $P$
with the \dmw\ test gives undersized tests with low power, so this
practice may favor simple benchmark models too much.  Existing
corrections, proposed by \citet{ClM:01,ClM:05}, \citet{Mcc:07} and
\citet{ClW:06,ClW:07}, seem to overcorrect for this problem, though, and reject
too often when the benchmark model is more accurate.

Although \oos\ comparisons have been popular in Macroeconomics and
Finance since \citepos{MeR:83} seminal study of exchange rate models,
it has been unclear from a theoretical perspective whether or not the
statistics are useful.  Empirical researchers often cite ``overfit''
or ``instability'' as reasons for using \oos\ comparisons, as in
\citet{StW:03}, but neither term is precisely defined or formalized.
Compounding this problem, the asymptotic distributions of these
statistics are derived under conditions that rule out either
instability or overfit and allow a researcher to use a conventional
in-sample comparison---a variation of the $F$-test, for example.  As
\citet{InK:04} argue, the statistics themselves are designed to test
hypotheses that can be tested by these in-sample statistics.  For
example, \citet{DiM:95} and \citet{Wes:96} derive the limiting
distributions of many popular \oos\ test statistics under conditions
that would justify these full-sample tests.  Much
of the subsequent research by \citet{Mcc:00, Mcc:07}, \citet{CCS:01},
\citet{ClM:01,ClM:05}, \citet{CoS:02,CoS:04}, \citet{ClW:06,ClW:07},
\citet{Ana:07}, and others relaxes several of Diebold and Mariano's
and West's assumptions, but maintains the stationarity and dependence
conditions that permit in-sample comparisons \citep[see][for a
review of this literature]{Wes:06}.\footnote{Like us, \citet{Ana:07}
  allows the number of regressors to increase with $T$.  But in that
  paper, the number of regressors increases slowly enough that the \ols\
  coefficients are consistent and asymptotically normal.} \citet{GiW:06} and
\citet{GiR:09, GiR:10} are exceptions.  Instead of focusing on
hypotheses that can be tested by in-sample comparisons, \citet{GiW:06}
derive an \oos\ test for the null hypothesis that the difference
between two models' \oos\ forecasting performance is unpredictable, a
martingale difference sequence; \citet{GiR:09} test whether the \oos\
forecasting performance of a model suffers from a breakdown relative
to its in-sample performance; and \citet{GiR:10} test whether the
forecasting performance is stable. However, those papers focus on a
particular \oos\ estimation strategy and do not address why \oos\
comparisons might be useful as a general strategy.

Since in-sample and \oos\ statistics require similar assumptions and
test similar hypotheses, one might expect that they would give similar
results.  They do not.  In-sample analyses tend to support more
complicated theoretical models and \oos\ analyses support simple
benchmarks, as seen in \citet{MeR:83}, \citet{StW:03}, and
\citet{GoW:08}.  Since these different approaches strongly influence
the outcome of research, it is important to know when each is
appropriate.  The explanations in favor of \oos\ comparisons claim
that they should be more robust to unmodeled instability
\citep{ClM:05,GiW:06,GiR:09,GiR:10} or to overfit
\citep{Mcc:98,Cla:04}.  Both explanations presume that the in-sample
comparison is invalid and the \oos\ comparisons are more reliable.  Of
course, as \citet{InK:04,InK:06} point out, both in-sample and \oos\
methods could be valid, but the \oos\ methods could have lower power.

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---the number of its regressors
increases with the number of observations so that their ratio remains
positive.  Although overfit is sometimes used to describe the
situation where a forecaster chooses from many different models,
i.e. \textit{data-mining} or \textit{data-snooping}, we view these as
separate issues.  Procedures that account for the presence of many
models have been, and are continuing to be, developed \citep[see, for
example,][]{Whi:00,Han:05,RoW:05,HHK:10,ClM:12b}, but it is unclear
whether those procedures should themselves use in-sample or \oos\
comparisons.  Understanding the difference between in-sample and \oos\
comparisons in the context of a simple comparison between two models
is necessary before resolving any new issues that arise with multiple
comparisons. Moreover, the empirical research that motivates
this paper uses \textit{pseudo} \oos\ comparisons and not
\textit{true} \oos\ comparisons.  Even if a true \oos\ comparison
could account for some forms of data-snooping better than
\citepos{Whi:00} \textsc{brc} or its extensions, in-sample and pseudo
\oos\ comparisons would both be affected by the data-snooping, a
point also made by \citet{InK:04}.

We focus on linear regression models estimated with a fixed window
for simplicity, but our basic
conclusions should be true for other models as well.  Under this
asymptotic theory, where the number of regressors $k$ increases with $T$
so that $\lim k/T$ is positive and less than one,
the \ols\ coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual \oos\ average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.}  Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudo-true
models, so existing in-sample and \oos\ comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true \dgp\ in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an \oos\ comparison would correctly fail to reject the null that the
benchmark is more accurate.\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and \oos\ tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the \dgp\ that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do \oos\ comparisons.}
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample even if there are no
structural breaks or other forms of instability.  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99} and
\citealp{StW:03} among others), but this breakdown can be caused by overfit as
well.

Our results can justify \oos\ comparisons when researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.  We show that the \dmw\ test
lets the forecaster control the probability of the first error, just
as with conventional hypothesis testing.

But we also identify new practical limitations for applying the \dmw\
test to overfit models.
Since the models' coefficients are imprecisely estimated in the limit,
the test sample must be small enough that the model estimated over the
training sample is similar to the one that will be estimated over the
full sample.  In particular, $P/T \to 0$ is required for consistent
estimation of the difference in the two models' performance,
and $P^2/T \to 0$ is required for valid confidence
intervals and inference.  For larger $P$, the \oos\ comparisons remain
asymptotically normal, but are centered on the forecasting performance
associated with the period-$R$ estimates.  In practice, researchers
typically use large values of $P$, so these studies may be too
pessimistic about their models' future accuracy if they use the \dmw\
test.  Section~\ref{sec:oostheory} lays out the asymptotic behavior of
the \dmw\ test under this limit theory.

Popular in-sample tests and model selection criteria, like the Wald
test, AIC, and BIC, do not help forecasters in this setting. We show
that these statistics can no pick out the more accurate model. For
many \dgp s they will choose a larger alternative model over a small
benchmark model with probability approaching one, regardless of which
model is more accurate.\footnote{Our result holds for a broad class of
  full-sample statistics, but there may be other potential statistics
  that mimic the \oos\ test and remain valid. Exploring such
  statistics is left for future research.}
This result holds even though modifications of the $F$-test are valid
under this asymptotic theory, as shown by \citet{BoB:95},
\citet{AkA:00}, \citet{AkP:04}, \citet{Cal:11c}, and \citet{Ana:12},
among others.\footnote{Also see \citet{Efr:86,Efr:04} for a discussion
  of naive in-sample loss comparisons.}
Moreover, under this asymptotic theory, many recent \oos\ test
statistics, such as those derived by \cite{ClM:01,ClM:05},
\citet{Mcc:07}, and \citet{ClW:06,ClW:07} should have the same
problems as in-sample tests.\footnote{Our theoretical results apply
  directly to \citepos{Mcc:07} \oos\ $t$-test, since it simply
  proposes more liberal critical values for the same test statistic
  that we study.  Since \citet{ClW:06,ClW:07} use a finite length
  estimation window, our asymptotics are incompatible with theirs and
  prevent us from studying their test directly, as well as
  \citepos{GiW:06} and other tests based on \citepos{GiW:06}
  asymptotics.  But \clws\ test can be viewed as a stochastic
  adjustment to the critical values of the usual \oos\ $t$ test, so
  our conclusions should apply informally as well.  Specifically, we
  show that the \dmw\ test rejects with probability equal to nominal
  size when the estimated benchmark model is expected to be more
  accurate, so more liberal critical values result in overrejection.}
These tests are also designed to reject the benchmark when the
alternative model is true, and so they may reject too often when the
benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the \dmw\ test when $P$ is small, but should overreject the benchmark
when $P$ is large. Section~\ref{sec:insample} presents our theoretical
results for full-sample statistics and Section~\ref{sec:mc} presents
Monte Carlo evidence to support these claims.

Finally, this paper also introduces a new method of proof for \oos\
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of \oos\ loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so 
the arguments used to prove asymptotic
results for \ned\ functions of mixing processes can be used for these
\oos\ processes with only slight modification.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
\dmw\ \oos\ test, shows that standard in-sample tests reject the
benchmark model too often when it is false but more accurate than the
alternative, and shows that standard model selection methods choose the
larger model too often under similar conditions. Section~\ref{sec:mc}
presents a Monte Carlo study supporting our theoretical results.
Section~\ref{sec:empirics} applies the \oos\ statistic to
\citepos{GoW:08} dataset for equity premium prediction, and
Section~\ref{sec:conclusion} concludes.  Proofs and supporting results
are listed in the Appendix.

@@

The models are then compared by their forecast errors over the test
sample.  There are many statistics that have been considered in the
literature, but we focus on perhaps the most natural, the \dmw\ \oost\
test \citep{DiM:95,Wes:96}.\footnote{The core insights of our paper
  apply to other \oos\ statistics as well.} This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation}
where
\begin{equation}\label{eq:16}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation}
and $L$ is a known loss function. The \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or \hac, estimator.)

@@

The conditional expectation $\E_T \oosB$ has been studied heavily in
cross-sectional settings with independent observations. In such a
setting, $\E_T \oosB$ is equal to the difference in the models'
\emph{generalization error}, which has been used widely as a measure
of model accuracy in the machine
learning literature \citep[see][for further discussion]{HTF:08}.
Moreover, with i.i.d. observations, the expectation of $\E_T \oosB$
equals \citepos{Aka:69} Final Prediction Error (\textsc{fpe}). Both
generalization error and \textsc{fpe} are defined by a model's
performance on a new, independent, data set, but, for lack of a better
term, we will call $\E_T \oosB$ the difference in generalization error
for the rest of the paper with hopefully no risk of confusion.

@@

The next conditions are assumed to hold throughout the paper.  The first
assumption controls the dependence of the underlying random array.
The second layes out the details of our asymptotic approximation.
The third assumption controls the
smoothness of the loss function and bounds the moments of the
difference in the models' performance; the fourth assumption describes
the behavior of the estimation and test windows.  And the last
assumption describes the kernel used to estimate the \oos\ average's
asymptotic variance.

@@

This assumption is a standard condition on the dependence of the
underlying stochastic array. The only novelty is that we use
absolute regularity instead of strong or uniform mixing as our
weak dependence condition; absolute regularity admits a paricular
coupling argument, ``Berbee's Lemma'' \citep[reproduced in this
paper as Lemma A.1 for reference]{Ber:79} that is
unavailable for strong mixing sequences; absolute regularity
implies uniform mixing, so this assumption is not unduly strong.
For a detailed discussion of these weak dependence conditions,
please see, for example, \citet{Dav:94}.

@@

The assumption that $y_{t+\h}$ and $x_t$ have positive and finite
variance is straightforward.  The conditions on the eigenvalues are
technical and control the behavior of the \ols\ estimator as the
number of regressors gets large.  The restrictions on the pseudo-true
coefficients ensure that the regression model doesn't dominate the
variance of $y_{t+\h}$ in the limit.

@@

  So the conditions amount to restrictions on the eigenvalues of
  $X_T'X_T$, which are familiar from general \ols.
\end{example}

The next assumption establishes moment conditions for the \oos\ loss
process and smoothness conditions for the loss function itself. The
moment conditions are standard and apply to $D_t$, and the smoothness
conditions are relatively weak.

@@

The differentiability condition in Assumption~\ref{asmp-3} is weak and
allows the loss function itself to be non-differentiable; for example,
absolute error and many asymmetric loss functions satisfy this
assumption.

Assumption~\ref{asmp-4} controls the growth of the test and training
samples.
\begin{asmp} \label{asmp-4} $P, R, Q \to\infty$ and $P^2/T \to 0$ as
  $T \to \infty$.
\end{asmp}

>>>>

\documentclass[11pt]{article}
\usepackage{tikz,setspace}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage[round]{natbib}
\usepackage[margin=1in]{geometry}
\onehalfspacing
\input{macros}
\usepackage[charter]{mathdesign}

@@

\title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: gcalhoun@iastate.edu. I
    would like to thank Julian Betts, Helle Bunzel, Marjorie Flavin,
    Nir Jaimovich, Lutz Kilian, Ivana Komunjer, Michael McCracken,
    Seth Pruitt, Ross Starr, Jim Stock, Yixiao Sun, Allan Timmermann,
    Hal White, several anonymous referees, participants at the 2010
    Midwest Economics Association Annual Meetings, the 2010
    International Symposium on Forecasting, the 2010 Joint Statistical
    Meetings, the 2011 NBER-NSF Time-Series
    Conference, and in many seminars at UCSD, and espeically
    Graham Elliott for their valuable suggestions, feedback and advice
    in writing this paper.  I would also like to thank Amit Goyal for
    providing computer code and data for his 2008 RFS paper
    with Ivo Welch \citep{GoW:08}.} \\ Iowa State University}

@@

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Our analysis gives a theoretical motivation for using
  out-of-sample (OOS) comparisons: the DMW OOS test allows a forecaster to
  conduct inference about the expected future accuracy of his or her
  models when one or both is overfit.  We show analytically and
  through Monte Carlo that standard full-sample test statistics can
  not test hypotheses about this performance.  Our paper also shows
  that popular test and training sample sizes may give misleading
  results if researchers are concerned about overfit.  We show that
  $P^2/T$ must converge to zero for the DMW test to give valid
  inference about the expected forecast accuracy, otherwise the test
  measures the accuracy of the estimates constructed using only the
  training sample.  In empirical research, $P$ is typically much
  larger than this.  Our simulations indicate that using large values
  of $P$ with the DMW test gives undersized tests with low power, so
  this practice may favor simple benchmark models too much.

\noindent JEL Classification: C12, C22, C52, C53

@@

\section{Introduction}\label{sec:introduction}
Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use OOS instead of in-sample
comparisons: the DMW OOS test\footnote{In this paper, we will
  refer to the basic OOS $t$-test studied by \citet{DiM:95} and
  \citet{Wes:96} as the DMW test.} allows a forecaster to conduct
inference about the expected future accuracy of his or her models when
one or both is overfit.  We show analytically and through Monte Carlo
that standard full-sample test statistics can not test hypotheses
about this performance.

Our paper also shows that popular test and training sample sizes may
give misleading results if researchers are concerned about overfit.
We show that $P^2/T$ must converge to zero for the DMW test to give
valid inference about the expected forecast accuracy, otherwise the
test measures the accuracy of the estimates constructed using only the
training sample.  In empirical research, $P$ is typically much larger
than this.  Our simulations indicate that using large values of $P$
with the DMW test gives undersized tests with low power, so this
practice may favor simple benchmark models too much.  Existing
corrections, proposed by \citet{ClM:01,ClM:05}, \citet{Mcc:07} and
\citet{ClW:06,ClW:07}, seem to overcorrect for this problem, though, and reject
too often when the benchmark model is more accurate.

Although OOS comparisons have been popular in Macroeconomics and
Finance since \citepos{MeR:83} seminal study of exchange rate models,
it has been unclear from a theoretical perspective whether or not the
statistics are useful.  Empirical researchers often cite ``overfit''
or ``instability'' as reasons for using OOS comparisons, as in
\citet{StW:03}, but neither term is precisely defined or formalized.
Compounding this problem, the asymptotic distributions of these
statistics are derived under conditions that rule out either
instability or overfit and allow a researcher to use a conventional
in-sample comparison---a variation of the $F$-test, for example.  As
\citet{InK:04} argue, the statistics themselves are designed to test
hypotheses that can be tested by these in-sample statistics.  For
example, \citet{DiM:95} and \citet{Wes:96} derive the limiting
distributions of many popular OOS test statistics under conditions
that would justify these full-sample tests.  Much
of the subsequent research by \citet{Mcc:00, Mcc:07}, \citet{CCS:01},
\citet{ClM:01,ClM:05}, \citet{CoS:02,CoS:04}, \citet{ClW:06,ClW:07},
\citet{Ana:07}, and others relaxes several of Diebold and Mariano's
and West's assumptions, but maintains the stationarity and dependence
conditions that permit in-sample comparisons \citep[see][for a
review of this literature]{Wes:06}.\footnote{Like us, \citet{Ana:07}
  allows the number of regressors to increase with $T$.  But in that
  paper, the number of regressors increases slowly enough that the OLS
  coefficients are consistent and asymptotically normal.} \citet{GiW:06} and
\citet{GiR:09, GiR:10} are exceptions.  Instead of focusing on
hypotheses that can be tested by in-sample comparisons, \citet{GiW:06}
derive an OOS test for the null hypothesis that the difference
between two models' OOS forecasting performance is unpredictable, a
martingale difference sequence; \citet{GiR:09} test whether the OOS
forecasting performance of a model suffers from a breakdown relative
to its in-sample performance; and \citet{GiR:10} test whether the
forecasting performance is stable. However, those papers focus on a
particular OOS estimation strategy and do not address why OOS
comparisons might be useful as a general strategy.

Since in-sample and OOS statistics require similar assumptions and
test similar hypotheses, one might expect that they would give similar
results.  They do not.  In-sample analyses tend to support more
complicated theoretical models and OOS analyses support simple
benchmarks, as seen in \citet{MeR:83}, \citet{StW:03}, and
\citet{GoW:08}.  Since these different approaches strongly influence
the outcome of research, it is important to know when each is
appropriate.  The explanations in favor of OOS comparisons claim
that they should be more robust to unmodeled instability
\citep{ClM:05,GiW:06,GiR:09,GiR:10} or to overfit
\citep{Mcc:98,Cla:04}.  Both explanations presume that the in-sample
comparison is invalid and the OOS comparisons are more reliable.  Of
course, as \citet{InK:04,InK:06} point out, both in-sample and OOS
methods could be valid, but the OOS methods could have lower power.

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of OOS comparisons when at least
one of the models is overfit---the number of its regressors
increases with the number of observations so that their ratio remains
positive.  Although overfit is sometimes used to describe the
situation where a forecaster chooses from many different models,
i.e. \textit{data-mining} or \textit{data-snooping}, we view these as
separate issues.  Procedures that account for the presence of many
models have been, and are continuing to be, developed \citep[see, for
example,][]{Whi:00,Han:05,RoW:05,HHK:10,ClM:12b}, but it is unclear
whether those procedures should themselves use in-sample or OOS
comparisons.  Understanding the difference between in-sample and OOS
comparisons in the context of a simple comparison between two models
is necessary before resolving any new issues that arise with multiple
comparisons. Moreover, the empirical research that motivates
this paper uses \textit{pseudo} OOS comparisons and not
\textit{true} OOS comparisons.  Even if a true OOS comparison
could account for some forms of data-snooping better than
\citepos{Whi:00} BRC or its extensions, in-sample and pseudo
OOS comparisons would both be affected by the data-snooping, a
point also made by \citet{InK:04}.

We focus on linear regression models estimated with a fixed window
for simplicity, but our basic
conclusions should be true for other models as well.  Under this
asymptotic theory, where the number of regressors $k$ increases with $T$
so that $\lim k/T$ is positive and less than one,
the OLS coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual OOS average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.}  Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudo-true
models, so existing in-sample and OOS comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true DGP in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an OOS comparison would correctly fail to reject the null that the
benchmark is more accurate.\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and OOS tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the DGP that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do OOS comparisons.}
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample even if there are no
structural breaks or other forms of instability.  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99} and
\citealp{StW:03} among others), but this breakdown can be caused by overfit as
well.

Our results can justify OOS comparisons when researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.  We show that the DMW test
lets the forecaster control the probability of the first error, just
as with conventional hypothesis testing.

But we also identify new practical limitations for applying the DMW
test to overfit models.
Since the models' coefficients are imprecisely estimated in the limit,
the test sample must be small enough that the model estimated over the
training sample is similar to the one that will be estimated over the
full sample.  In particular, $P/T \to 0$ is required for consistent
estimation of the difference in the two models' performance,
and $P^2/T \to 0$ is required for valid confidence
intervals and inference.  For larger $P$, the OOS comparisons remain
asymptotically normal, but are centered on the forecasting performance
associated with the period-$R$ estimates.  In practice, researchers
typically use large values of $P$, so these studies may be too
pessimistic about their models' future accuracy if they use the DMW
test.  Section~\ref{sec:oostheory} lays out the asymptotic behavior of
the DMW test under this limit theory.

Popular in-sample tests and model selection criteria, like the Wald
test, AIC, and BIC, do not help forecasters in this setting. We show
that these statistics can no pick out the more accurate model. For
many DGPs they will choose a larger alternative model over a small
benchmark model with probability approaching one, regardless of which
model is more accurate.\footnote{Our result holds for a broad class of
  full-sample statistics, but there may be other potential statistics
  that mimic the OOS test and remain valid. Exploring such
  statistics is left for future research.}
This result holds even though modifications of the $F$-test are valid
under this asymptotic theory, as shown by \citet{BoB:95},
\citet{AkA:00}, \citet{AkP:04}, \citet{Cal:11c}, and \citet{Ana:12},
among others.\footnote{Also see \citet{Efr:86,Efr:04} for a discussion
  of naive in-sample loss comparisons.}
Moreover, under this asymptotic theory, many recent OOS test
statistics, such as those derived by \cite{ClM:01,ClM:05},
\citet{Mcc:07}, and \citet{ClW:06,ClW:07} should have the same
problems as in-sample tests.\footnote{Our theoretical results apply
  directly to \citepos{Mcc:07} OOS $t$-test, since it simply
  proposes more liberal critical values for the same test statistic
  that we study.  Since \citet{ClW:06,ClW:07} use a finite length
  estimation window, our asymptotics are incompatible with theirs and
  prevent us from studying their test directly, as well as
  \citepos{GiW:06} and other tests based on \citepos{GiW:06}
  asymptotics.  But \clws\ test can be viewed as a stochastic
  adjustment to the critical values of the usual OOS $t$ test, so
  our conclusions should apply informally as well.  Specifically, we
  show that the DMW test rejects with probability equal to nominal
  size when the estimated benchmark model is expected to be more
  accurate, so more liberal critical values result in overrejection.}
These tests are also designed to reject the benchmark when the
alternative model is true, and so they may reject too often when the
benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the DMW test when $P$ is small, but should overreject the benchmark
when $P$ is large. Section~\ref{sec:insample} presents our theoretical
results for full-sample statistics and Section~\ref{sec:mc} presents
Monte Carlo evidence to support these claims.

Finally, this paper also introduces a new method of proof for OOS
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of OOS loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so 
the arguments used to prove asymptotic
results for Near Epoch Dependent (NED) functions of mixing processes
can be used for these OOS processes with only slight modification.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
DMW OOS test, shows that standard in-sample tests reject the
benchmark model too often when it is false but more accurate than the
alternative, and shows that standard model selection methods choose the
larger model too often under similar conditions. Section~\ref{sec:mc}
presents a Monte Carlo study supporting our theoretical results.
Section~\ref{sec:empirics} applies the OOS statistic to
\citepos{GoW:08} dataset for equity premium prediction, and
Section~\ref{sec:conclusion} concludes.  Proofs and supporting results
are listed in the Appendix.

@@

The models are then compared by their forecast errors over the test
sample.  There are many statistics that have been considered in the
literature, but we focus on perhaps the most natural, the DMW \oost\
test \citep{DiM:95,Wes:96}.\footnote{The core insights of our paper
  apply to other OOS statistics as well.} This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation}
where
\begin{equation}\label{eq:16}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation}
and $L$ is a known loss function. The \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or HAC, estimator.)

@@

The conditional expectation $\E_T \oosB$ has been studied heavily in
cross-sectional settings with independent observations. In such a
setting, $\E_T \oosB$ is equal to the difference in the models'
\emph{generalization error}, which has been used widely as a measure
of model accuracy in the machine
learning literature \citep[see][for further discussion]{HTF:08}.
Moreover, with i.i.d. observations, the expectation of $\E_T \oosB$
equals \citepos{Aka:69} Final Prediction Error (FPE). Both
generalization error and FPE are defined by a model's
performance on a new, independent, data set, but, for lack of a better
term, we will call $\E_T \oosB$ the difference in generalization error
for the rest of the paper with hopefully no risk of confusion.

@@

The next conditions are assumed to hold throughout the paper.  The first
assumption controls the dependence of the underlying random array.
The second lays out the details of our asymptotic approximation.
The third assumption controls the
smoothness of the loss function and bounds the moments of the
difference in the models' performance; the fourth assumption describes
the behavior of the estimation and test windows.  And the last
assumption describes the kernel used to estimate the OOS average's
asymptotic variance.

@@

This assumption is a standard condition on the dependence of the
underlying stochastic array. The only novelty is that we use
absolute regularity instead of strong or uniform mixing as our
weak dependence condition; absolute regularity admits a particular
coupling argument, ``Berbee's Lemma'' \citep[reproduced in this
paper as Lemma A.1 for reference]{Ber:79} that is
unavailable for strong mixing sequences; absolute regularity
implies uniform mixing, so this assumption is not unduly strong.
For a detailed discussion of these weak dependence conditions,
please see, for example, \citet{Dav:94}.

@@

The assumption that $y_{t+\h}$ and $x_t$ have positive and finite
variance is straightforward.  The conditions on the eigenvalues are
technical and control the behavior of the OLS estimator as the
number of regressors gets large.  The restrictions on the pseudo-true
coefficients ensure that the regression model doesn't dominate the
variance of $y_{t+\h}$ in the limit.

@@

  So the conditions amount to restrictions on the eigenvalues of
  $X_T'X_T$, which are familiar from general OLS.
\end{example}

The next assumption establishes moment conditions for the OOS loss
process and smoothness conditions for the loss function itself. The
moment conditions are standard and apply to $D_t$, and the smoothness
conditions are relatively weak.

@@

The differentiability condition in Assumption~\ref{asmp-3} is weak and
allows the loss function itself to be non-differentiable; for example,
absolute error and many asymmetric loss functions satisfy this
assumption.  The next assumption controls the growth of the test and
training samples.
\begin{asmp} \label{asmp-4} $P, R, Q \to\infty$ and $P^2/T \to 0$ as
  $T \to \infty$.
\end{asmp}
